A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-05-08 paper/paper.tex
paper: minor formatting fixes

% Simplicial and Cubical sets
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{{\Box}}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\Box}}}}

@@

The idea for such indexed presentation of presheaves over a direct category was introduced at the Univalent Foundations year in the context of defining semi-simplicial types~\cite{}. A couple of constructions were proposed since then. One, by Voevodsky~\cite{Voevodsky12}, relies on the presentation of semi-simplicial sets as a presheaf over injective monotone maps between finite ordinals. Another, by the first author, and fully formalized in Coq, relies on the presentation of semi-simplicial sets as a presheaf over face maps~\cite{herbelin15}. Another one by Part and Luo, and formalized in the Plastic proof assistant, works generically for any Reedy category with only upwards maps~\cite{part15}.

@@

\subsection{\texorpdfstring{$\nu$}{ν}-sets}
Let us call $\nu$-sets, the generalization of augmented semi-simplicial sets and semi-cubical sets. To obtain this, we extend $\DeltaPlus$ and $\Cube$ in a straightforward manner into a category which we call $\nu$-semi-shape category. The morphisms of the $\nu$-semi-shape category are sequences of elements of a set $\nu$ of arbitrary cardinal, extended with $\kstar$, so that the following holds.

\begin{table}[H]
  \centering
  \begin{tabularx}{0.95\linewidth}{X|X|X}
    \toprule
    Value of $\nu$ & 1                               & 2                  \\
    \graymidrule
    Interpretation & Augmented semi-simplicial types & Semi-cubical types \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\subsection{Augmented semi-simplicial sets}
Augmented semi-simplicial sets are defined similarly to semi-simplicial sets, except that the connected components are additionally dependent on a ``color''. Conversely, semi-simplicial sets can be seen as augmented semi-simplicial sets over a singleton set of colors.

@@

\subsection{Semi-cubical sets}
Semi-cubical sets are defined like augmented semi-simplicial sets except that $\DeltaPlus$ is replaced by $\Cube$ in which we take sequences of $L$, $R$ and $\star$, instead of sequences of $0$ and $\kstar$. They represent ordinary semi-cubical sets, with faces (and no augmentation).

@@

\section{Comparison of fibered and indexed presentations}
%     - standard augmented triangle:
%       Inductive points := A | B | C.
%       Inductive lines := AB | BC | AC.
%       Inductive surface := ABC.

@@

\section{Our construction}

Our construction is highly dependent, with dependencies in inequality
and equality proofs. It involves several involved details, some of
them being visible only in the Coq formalization.

@@

\subsection{Formal construction in ETT\label{sec:ett}}
% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{painting}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrpainting}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohpainting}[D, d]

@@

\subsection{Intuition for our formal construction\label{sec:intuition}}
To assign types to $X_0, X_1, X_2, \ldots$ in the above indexed representation, we give a recursive definition relying on building blocks, which we call $\framep$, $\layer$, and $\painting$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), which we decompose into $\layer$, and a $\painting$ corresponds to a filled $\framep$. Some $\framep$ are full and we call them $\fullframe$.

@@

\section{Details on the mechanization\label{sec:mech}}
\subsection{Formal construction in ITT\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\renewcommandx{\cohl}{\coh{layer}[][][][][][][][]}

@@

\subsection{Finer details}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][#1][][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][#1][][]}
\renewcommandx{\cohc}[1][1]{\coh{painting}[][][][][][#1][][]}

@@

Another direction for future work would also be to generically construct the indexed form of any presentation of presheaves over a direct category and to show the equivalence between the two presentations.
\bibliographystyle{msclike}
\bibliography{paper}

>>>>

% Simplicial and Cubical sets
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{{\square}}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

@@

The idea for such indexed presentation of presheaves over a direct category was introduced at the Univalent Foundations year in the context of defining semi-simplicial types~\cite{}. A couple of constructions were proposed since then. One, by Voevodsky~\cite{voevodsky12}, relies on the presentation of semi-simplicial sets as a presheaf over increasing injective maps between finite ordinals. Another, by the first author, and fully formalized in Coq, relies on the presentation of semi-simplicial sets as a presheaf over face maps~\cite{herbelin15}. Another one by Part and Luo, and formalized in the Plastic proof assistant, works generically for any Reedy category with only upwards maps~\cite{part15}.

@@

\subsection*{\texorpdfstring{$\nu$}{ν}-sets}
Let us call $\nu$-sets, the generalization of augmented semi-simplicial sets and semi-cubical sets. To obtain this, we extend $\DeltaPlus$ and $\Cube$ in a straightforward manner into a category which we call $\nu$-semi-shape category. The morphisms of the $\nu$-semi-shape category are sequences of elements of a set $\nu$ of arbitrary cardinal, extended with $\kstar$, so that the following holds.

\begin{table}[H]
  \centering
  \begin{tabularx}{0.7\linewidth}{X|X|X}
    \toprule
    Value of $\nu$ & 1                               & 2                  \\
    \graymidrule
    Interpretation & Augmented semi-simplicial types & Semi-cubical types \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\subsection*{Augmented semi-simplicial sets}
Augmented semi-simplicial sets are defined similarly to semi-simplicial sets, except that the connected components are additionally dependent on a ``color''. Conversely, semi-simplicial sets can be seen as augmented semi-simplicial sets over a singleton set of colors.

@@

\subsection*{Semi-cubical sets}
Semi-cubical sets are defined like augmented semi-simplicial sets except that $\DeltaPlus$ is replaced by $\Cube$ in which we take sequences of $L$, $R$ and $\star$, instead of sequences of $0$ and $\kstar$. They represent ordinary semi-cubical sets, with faces (and no augmentation).

@@

\section*{Comparison of fibered and indexed presentations}
%     - standard augmented triangle:
%       Inductive points := A | B | C.
%       Inductive lines := AB | BC | AC.
%       Inductive surface := ABC.

@@

\section{Our construction}
Our construction is highly dependent, with dependencies in inequality and equality proofs. It involves several involved details, some of them being visible only in the Coq formalization.

@@

\subsection*{Formal construction in ETT\label{sec:ett}}
% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{painting}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrpainting}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohpainting}[D, d]

@@

\subsection*{Intuition for our formal construction\label{sec:intuition}}
To assign types to $X_0, X_1, X_2, \ldots$ in the above indexed representation, we give a recursive definition relying on building blocks, which we call $\framep$, $\layer$, and $\painting$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), which we decompose into $\layer$, and a $\painting$ corresponds to a filled $\framep$. Some $\framep$ are full and we call them $\fullframe$.

@@

\section{Details on the mechanization\label{sec:mech}}
\subsection*{Formal construction in ITT\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\renewcommandx{\cohl}{\coh{layer}[][][][][][][][]}

@@

\subsection*{Finer details}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][#1][][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][#1][][]}
\renewcommandx{\cohc}[1][1]{\coh{painting}[][][][][][#1][][]}

@@

Another direction for future work would also be to generically construct the indexed form of any presentation of presheaves over a direct category and to show the equivalence between the two presentations.

\bibliographystyle{msclike}
\bibliography{paper}
